Bounds on Coding Theory\\from Algebraic Geometry

### 2021-05-06 draft.tex
2021-05-06 afternoon review update

Suppose Alice wants to send a message to Bob though a noisy channel.
When Bob receives the message, it is possible that some of the information is misinterpreted.
For example, if Alice sends the message `\texttt{food}' through the noisy channel, one of its letters may be misinterpreted and Bob could actually receive `\texttt{\underline{m}ood}'.
So, the question Coding Theory tries to answer is:
How can Alice and Bob agree on a system beforehand so that, if Alice sends a message to Bob, even if some misinterpretations occur, Bob will be able to understand the correct meaning?

@@

We will provide the proof for Theorem \ref{thm:singleton_bound} when $C$ is a linear code over a field $K$.

@@

\begin{defn}[Reed--Solomon codes, \cite{LS87}]\label{defn:rs_codes}
	Let $q$ be a power of a prime, and $\mathbb{F}_q = \{\alpha_1, \alpha_2, \dots, \alpha_q\}$ the field with $q$ elements.
	Let $k$ be an integer, and ${L}_k$ the set of all polynomials over $\mathbb{F}_q$ with degree smaller than $k$.
	Let $k \le n \le q$ be an integer. The Reed--Solomon code $RS_q(n, k)$ over $\mathbb{F}_q$ is 
	\begin{equation}
		RS_q(n, k) = \left\{\left(f(\alpha_1), f(\alpha_2), \dots, f(\alpha_n)\right) \mid f \in {L}_k\right\}.
	\end{equation}
\end{defn}

@@

\begin{proof}
	$RS_q(n, k)$ is a subset of $\mathbb{F}_q^n$, thus it has length $n$.  
	Note that $ \mathcal L_k$ is a vector space over $\mathbb{F}_q$.
	Note that $\{1, x, x^2, \dots, x^{k-1}\}$ is a choice of basis for this vector space, thus it has dimension $k$.
	Consider the map $\phi:  \mathcal L_k \to \mathbb{F}_q^n$ given by 
	\begin{equation}
		f \mapsto (f(\alpha_1), f(\alpha_2), \dots, f(\alpha_n)).
	\end{equation}

	Note that the map $\phi$ is a linear transformation.
	Thus, its image $\Ima \phi = RS_q(n, k)$ is also a vector space.  
	Additionally, if $\phi(f) = \phi(g)$, then $f - g$ has at least $n$ roots, but has degree less than $n$; thus $f - g$ is the zero polynomial, which implies $f = g$.
	Therefore, $\phi$ is also injective.
	This implies that the dimension of the domain $ \mathcal L_k$ is the same as the dimension of the image $RS_q(n, k)$, i.e.,  $\dimension RS_q(n, k) = k$.

	Finally, consider distinct $f, g \in  \mathcal L_k$ and define $d = \dist(\phi(f), \phi(g))$, $f - g$ has at least $n - d$ roots.
	Furthermore, $f - g$ is a non-zero polynomial with degree less than $k$, thus has at most $k - 1$ roots.
	Then,  $ k - 1 \ge \#\ \text{roots} \ge n - d$.
	If we choose $f, g$ such that $d$ is the minimal distance, we get $k + d_{\mathrm{min}} \ge n + 1$, which together with the Singleton Bound, implies $k + d_\text{min} = n + 1$.
\end{proof}

The Reed--Solomon codes are very good codes in the sense that they have the largest possible sum $k + d$ for their length $n$.
However, Reed--Solomon codes are limited because their length is at most the alphabet size.
So, a question naturally arises: Given fixed $\mathbb{F}_q$, are there codes over $\mathbb{F}_q$ with arbitrarily large $n$ and $R + \delta = 1 + 1/n$?
If not, how large can $R$ and $\delta$ be when $n$ gets larger?
The well-known Gilbert--Varshamov bound shows that there are codes with 
\begin{equation}
	1 - R \approx H_q(\delta), \text{ as $n \to \infty$,}
\end{equation}
in which
\begin{equation}
	H_q(x) = x \log_q(q-1) - x\log_q(x) - (1-x)\log_q(x).
\end{equation}
The article \cite{TVZ82} uses Algebraic Geometry to show that there are codes that give better bounds.

@@

Through this section, let $K$ be a field and $L$ its algebraic closure.

@@

Note that there exists a natural injection, called \emph{homogenization}, from the set of polynomials in $K[x, y]$ and the set of homogeneous polynomials in $K[X, Y, Z]$ defined by $f(x, y) \mapsto Z^{\deg f}f(X/Z, Y/Z)$.
Because of this, we'll denote by $\mathfrak C$ both $\mathfrak C_f$ and $\mathfrak C_F$, where the usage is clear by context.

@@

\begin{exmp*}
	Consider the curve $\mathfrak C$ defined by $X^2 + Y^2 + Z^2 = 0$ over $\mathbb{F}_7$.
	The point $(1 : 2 : 3)$ is a $\mathbb{F}_7$-rational point on $\mathfrak C$.
	Note that the polynomial $x^2 - 3$ is irreducible over $\mathbb{F}_7$, but there it has a root $\alpha$ in its algebraic closure, $\overline{\mathbb{F}_7}$. The projective point $(1 :  \alpha : \alpha)$ is a $\overline{\mathbb{F}_7}$-rational point on $\mathfrak C$.
\end{exmp*}

@@

\begin{defn}[Rational function on a curve]
	Let $F \in K[X, Y, Z]$ be a homogeneous polynomial and let $\mathfrak C = \mathfrak C_F$.
	If $G, H \in K[X, Y, Z]$ are homogeneous polynomials of equal degree, then the fraction $G/H$ is called a \emph{rational function} on $\mathfrak C$.
	Fractions $G/H$ and $G'/H'$ define the same rational function if $G'H - GH'$ is a multiple of $F$, i.e., if  $G'H - GH'$ vanishes at all $L$-rational points on $\mathfrak C$.
\end{defn}

	Note that it makes sense to evaluate a rational function on a projective point, since \[
		\frac{G(X_0, Y_0, Z_0)}{H(X_0, Y_0, Z_0)} = \frac{\lambda^d G(X_0, Y_0, Z_0)}{\lambda^d H(X_0, Y_0, Z_0)} = \frac{G(\lambda X_0, \lambda Y_0, \lambda Z_0)}{H(\lambda X_0, \lambda Y_0, \lambda Z_0)},
	\]
	i.e., the result is the same regardless of the representation of a given projective point.

	After cancelling common factors of $G$ and $H$, a $L$-rational point $P$ on $\mathfrak C$ is called a \emph{zero} of $G/H$ whenever $G$ vanishes at $P$, and is called a \emph{poles} of $G/H$ whenever $H$ vanishes at $P$.  
	We say $G/H$ has order $n$ at $P$ if $P$ is a zero of multiplicity $n$; order $-n$ if $P$ is a pole with multiplicity $n$; and order $0$, otherwise.

\begin{defn}[Divisor] 
%	Let $\overline{\mathbb{F}_q}$ be the algebraic closure of $\mathbb{F}_q$.
%	Let $\mathfrak C$ be a projective plane curve defined over $\mathbb{F}_q$.
	Let $\mathfrak C$ be a projective plane curve defined over $K$.
	A \emph{divisor} $D$ on $\mathfrak C$ is a formal finite sum of the form $D = \sum a_P P$, where $P$ varies over the $L$-rational points on $\mathfrak C$, $a_P$ is an integer and $a_P = 0$ for all but finitely many points $P$.
	The \emph{degree} of $D$ is $\sum a_P$.
	The \emph{support} of $D$, denoted by $\Supp D$, is the set of $L$-rational points $P$ on $\mathfrak C$ that satisfy $a_P \neq 0$.
\end{defn}

@@

\begin{defn}[\cite{LS87}]
	If $D = \sum n_p P$, then define $\mathcal L(D)$ as the vector space formed by all homogeneous rational functions $f$ such that the order of $f$ at each $L$-rational point $P$ on $\mathfrak C$ is greater or equal to $-n_P$ (plus the zero polynomial).
\end{defn}

@@

\begin{proof}
	Bezout's theorem implies that $\mathfrak C$ and $\mathfrak C_G$ intersect in the same number of points, counting multiplicity, than $\mathfrak C$ and $\mathfrak C_H$. Therefore, the number of zeros of a function is the same as the number of poles, counting multiplicity. Thus, the sum of the orders of $G/H$ at the $L$-rational points on $\mathfrak C$ must be zero. On the other hand, if $G/H \in \mathcal L(D)$, then the sum of the orders of $G/H$ at the $L$-rational points on $\mathfrak C$ must be $\ge - \deg(D)$.

@@

\section{Generalized Reed--Solomon codes} \label{s:grs}
Let $\mathbb{P}^1(\mathbb{F}_q)$ denote the projective line over $\mathbb{F}_q = \{\alpha_1, \dots, \alpha_q\}$.
We will write $(a : b)$ to denote the projective point corresponding to the $1$-dimensional vector space containing $(a, b)$.
The points on $\mathbb{P}^1(\mathbb{F}_q)$ are the points
\begin{equation}
	P_i = (\alpha_i : 1), \quad 1 \le i \le q, \qquad \text{and} \qquad 
	P_\infty = (1 : 0).
\end{equation}

Consider the divisor $D = (k-1)P_\infty$ and the associated vector space $\mathcal L(D)$, which can be seen as the set of two-variable homogeneous rational functions which have a pole of order less than $k$ in the point $P_\infty$.

Note that, if a polynomial $f(x) \in \mathcal L_k$ has degree $d < k$, then the rational function \[
	F(X, Y) = \frac{Y^d f(X/Y)}{Y^d}
\]
has $P_\infty$ as its only pole, with order at most $k-1$. Thus, $F(X, Y) \in \mathcal L(D)$ and $f(\alpha_i) = F(P_i)$ for any $\alpha_i$.

Then, we can rewrite the Reed--Solomon code from Definition \ref{defn:rs_codes} as
\begin{equation}
	RS_q(n, k) = \left\{\big(f(P_1), f(P_2), \dots, f(P_n)\big) \mid f \in \mathcal L(D)\right\}.
\end{equation}

@@

\begin{thm}[\cite{Wal00}]
	Let $\mathfrak C, \mathcal P, D$ be as above. Let $g$ denote the genus of $\mathfrak C$. Suppose $D$ satisfies $2g - 2 < \deg D < n$. Then the algebraic geometric code $C(\mathfrak C, \mathcal P, D)$ is linear of:
	\begin{enumerate}[label = \textbullet, itemsep = 0pt]
		\item length $n$,
		\item dimension $\deg D + 1 - g$,
		\item minimum distance $d \ge n - \deg D$.
	\end{enumerate}
\end{thm}

@@

This theorem is an interesting result! We want large $R$ and $\delta$, and these codes yield $k + d \ge n + 1 - g$, i.e., 
	\begin{equation}
		\label{eq:g/n} R + \delta \ge 1 + 1/n - g/n,
	\end{equation}
where $n$ is the number of rational points of a curve $\mathfrak C$, with genus $g$.
Futhermore, all the work done can be applied for algebraic curves in higher dimensional projective spaces.

@@

On equation \eqref{eq:g/n}, we observe that good algebraic geometric codes are generated by curves with a small ratio between $g$ and $n$.
On \cite{TVZ82},\todo{Write one or two more sentences about this article.} the authors present a sequence of such curves, with $g/n$ large enough to create a better bound than the Gilbert--Varshamov one.

>>>>

Suppose Alice wants to send a message to Bob through a noisy channel.
When Bob receives the message, it is possible that some of the information is distorted.
For example, if Alice sends the message `\texttt{I am sad!}' through the noisy channel, one of its letters may be misinterpreted and Bob could actually receive the message `\texttt{I am \underline{m}ad!}'.
So, the question Coding Theory tries to answer is:
How can Alice and Bob agree on a system beforehand so that, if Alice sends a message to Bob, even if some misinterpretations occur, Bob will be able to understand the correct meaning?

@@

We will provide the proof for Theorem \ref{thm:singleton_bound} when $C$ is a linear code over a field $K$, but bear in mind that the theorem is true for any code $C$.

@@

\begin{defn}[Reed--Solomon codes, \cite{LS87}]\label{defn:rs_codes}
	Let $q$ be a power of a prime, and $\mathbb{F}_q = \{\alpha_1, \alpha_2, \dots, \alpha_q\}$ the field with $q$ elements.
	Let $k$ be a positive integer, and $\mathcal L_k$ the set of all polynomials over $\mathbb{F}_q$ with degree smaller than $k$.
	Let $k \le n \le q$ be an integer. The Reed--Solomon code $RS_q(n, k)$ over $\mathbb{F}_q$ is 
	\begin{equation}
		RS_q(n, k) = \left\{\left(f(\alpha_1), f(\alpha_2), \dots, f(\alpha_n)\right) \mid f \in \mathcal L_k\right\} \subset \mathbb{F}_q^n.
	\end{equation}
\end{defn}

@@

\begin{proof}
	$RS_q(n, k)$ is a subset of $\mathbb{F}_q^n$, thus it is a code with length $n$.  
	Note that $ \mathcal L_k$ is a vector space over $\mathbb{F}_q$, and $(1, x, x^2, \dots, x^{k-1})$ is a choice of basis for this vector space, thus it has dimension $k$.
	Consider the map $\phi:  \mathcal L_k \to \mathbb{F}_q^n$ given by 
	\begin{equation}
		f \mapsto (f(\alpha_1), f(\alpha_2), \dots, f(\alpha_n)),
	\end{equation}
	which is a linear transformation.
	Thus, its image $\Ima \phi = RS_q(n, k)$ is also a vector space.  
	Additionally, if $\phi(f) = \phi(g)$, then $f - g$ has at least $n$ roots, but has degree less than $n$; thus $f - g$ is the zero polynomial, which implies $f = g$.
	Therefore, $\phi$ is also injective.
	This implies that the dimension of the domain $ \mathcal L_k$ is the same as the dimension of the image $RS_q(n, k)$, i.e.,  $\dimension RS_q(n, k) = k$.

	Finally, consider $f \neq g \in  \mathcal L_k$ and define $d = \dist(\phi(f), \phi(g))$, $f - g$ has at least $n - d$ roots.
	Furthermore, $f - g$ is a non-zero polynomial with degree less than $k$, thus has at most $k - 1$ roots.
	Then,  $ k - 1 \ge \#\ \text{roots} \ge n - d$.
	If we choose $f$ and $g$ such that $d$ is the minimal distance, we get $k + d_{\mathrm{min}} \ge n + 1$, which together with the Singleton Bound, implies $k + d_\text{min} = n + 1$.
\end{proof}

The Reed--Solomon codes are very good codes in the sense that they have the largest possible sum $k + d$ for their length.
This is guaranteed by the Singleton Bound.

However, Reed--Solomon codes are limited because their length is at most their alphabet size.
So, a question naturally arises: Given fixed $\mathbb{F}_q$, are there codes over $\mathbb{F}_q$ with arbitrarily large $n$ and $R + \delta = 1 + 1/n$?
If not, how large can $R$ and $\delta$ be when $n$ gets larger?
The well-known Gilbert--Varshamov bound shows that there are codes with 
\begin{equation}
	1 - R \approx H_q(\delta), \text{ as $n \to \infty$,}
\end{equation}
in which
\begin{equation}
	H_q(x) = x \log_q(q-1) - x\log_q(x) - (1-x)\log_q(x).
\end{equation}
However, the article \cite{TVZ82} uses structures from Algebraic Geometry to create new codes and show that they (sometimes) give better bounds.

@@

Before understanding the codes from \cite{TVZ82}, we need to familiarize with some tools in Algebraic Geometry.
Through this section, let $K$ be a field and $L$ its algebraic closure.

@@

Note that there exists a natural injection, called \emph{homogenization}, from the set of polynomials in $K[x, y]$ to the set of homogeneous polynomials in $K[X, Y, Z]$ defined by $f(x, y) \mapsto Z^{\deg f}f(X/Z, Y/Z)$.
Because of this, we'll denote both $\mathfrak C_f$ and $\mathfrak C_F$ by $\mathfrak C$, where the usage is clear by context.

@@

\begin{exmp*}
	Consider the curve $\mathfrak C$ defined by $X^2 + Y^2 + Z^2 = 0$ over $\mathbb{F}_7$.
	The point $(1 : 2 : 3)$ is a $\mathbb{F}_7$-rational point on $\mathfrak C$.
	Note that the polynomial $x^2 - 3$ is irreducible over $\mathbb{F}_7$, but there is a root $\alpha$ in its algebraic closure, $\overline{\mathbb{F}_7}$. The projective point $(1 :  \alpha : \alpha)$ is a $\overline{\mathbb{F}_7}$-rational point on $\mathfrak C$.
\end{exmp*}

One of the most important theorems in Algebraic Geometry, that we will use in this paper, is the following one:

@@

\begin{defn}[Rational function on a curve, \cite{TVN07}]
	Let $F \in K[X, Y, Z]$ be a homogeneous polynomial and let $\mathfrak C = \mathfrak C_F$.
	If $G, H \in K[X, Y, Z]$ are homogeneous polynomials of equal degree, then the fraction $G/H$ is called a \emph{rational function} on $\mathfrak C$.
	Fractions $G/H$ and $G'/H'$ define the same rational function if $G'H - GH'$ is a multiple of $F$, i.e., if  $G'H - GH'$ vanishes at all $L$-rational points on $\mathfrak C$.
\end{defn}

	Note that it makes sense to evaluate a rational function on a projective point, since the result is the same regardless of the representation of the projective point, i.e.,  \[
		\frac{G(X_0, Y_0, Z_0)}{H(X_0, Y_0, Z_0)} = \frac{\lambda^d G(X_0, Y_0, Z_0)}{\lambda^d H(X_0, Y_0, Z_0)} = \frac{G(\lambda X_0, \lambda Y_0, \lambda Z_0)}{H(\lambda X_0, \lambda Y_0, \lambda Z_0)}.
	\]

	After cancelling common factors of $G$ and $H$, a $L$-rational point $P$ on $\mathfrak C$ is called a \emph{zero} of $G/H$ whenever $G$ vanishes at $P$, and is called a \emph{pole} of $G/H$ whenever $H$ vanishes at $P$.  
	We say $G/H$ has order $n$ at $P$ if $P$ is a zero of multiplicity $n$; order $-n$ if $P$ is a pole with multiplicity $n$; and order $0$, otherwise.

\begin{defn}[Divisor] 
%	Let $\overline{\mathbb{F}_q}$ be the algebraic closure of $\mathbb{F}_q$.
%	Let $\mathfrak C$ be a projective plane curve defined over $\mathbb{F}_q$.
	Let $\mathfrak C$ be a projective plane curve defined over $K$.
	A \emph{divisor} $D$ on $\mathfrak C$ is a formal finite sum of the form $D = \sum a_P P$, where the sum varies over all $L$-rational points $P$ on $\mathfrak C$, $a_P$ is an integer and $a_P = 0$ for all but finitely many $L$-rational points $P$ in $\mathfrak C$.

	The \emph{degree} of $D$ is $\sum a_P$.
	The \emph{support} of $D$, denoted by $\Supp D$, is the set of $L$-rational points $P$ on $\mathfrak C$ that satisfy $a_P \neq 0$.
\end{defn}

@@

\begin{defn}[\cite{LS87}]
	If $D = \sum a_p P$, then define $\mathcal L(D)$ as the vector space formed by all homogeneous rational functions $f$ such that the order of $f$ at each $L$-rational point $P$ on $\mathfrak C$ is greater or equal to $-a_P$ (plus the zero polynomial).
\end{defn}

@@

\begin{proof}
	Bezout's theorem implies that $\mathfrak C$ and $\mathfrak C_G$ intersect in the same number of points, counting multiplicity, than $\mathfrak C$ and $\mathfrak C_H$. Therefore, the number of zeros of a rational function $G/H$ is the same as the number of poles, counting multiplicity. Thus, the sum of the orders of $G/H$ at the $L$-rational points on $\mathfrak C$ must be zero. On the other hand, if $G/H \in \mathcal L(D)$, then the sum of the orders of $G/H$ at the $L$-rational points on $\mathfrak C$ must be $\ge - \deg(D)$.

@@

\section{Generalized Reed--Solomon codes} \label{s:grs}
Let's now contextualize Reed--Solomon codes inside Algebraic Geometry.

Let $\mathbb{P}^1(\mathbb{F}_q)$ be the projective line over $\mathbb{F}_q = \{\alpha_1, \dots, \alpha_q\}$.
We will write $(a : b)$ to denote the projective point corresponding to the $1$-dimensional vector space containing $(a, b)$.
Thus, the points on $\mathbb{P}^1(\mathbb{F}_q)$ are
\begin{equation}
	P_i = (\alpha_i : 1), \quad 1 \le i \le q, \qquad \text{and} \qquad 
	P_\infty = (1 : 0).
\end{equation}

Consider the divisor $D = (k-1)P_\infty$ and the associated vector space $\mathcal L(D)$, which can be understood as the set of two-variable homogeneous rational functions which can only have a pole at $P_\infty$, and with order at most $k-1$.

Note that, if a polynomial $f(x) \in \mathcal L_k$ has degree $d < k$, then the rational function \[
	F(X, Y) = \frac{Y^d f(X/Y)}{Y^d}
\]
has $P_\infty$ as its only pole, with order at most $k-1$. Thus, $F(X, Y) \in \mathcal L((k-1)P_\infty)$ and $f(\alpha_i) = F(P_i)$ for any $\alpha_i$.

On the other hand, any $G/H \in \mathcal L((k-1)P_\infty)$ has $H = Y^d$ with $d < k$ (since the only pole is $P_\infty$) and, therefore, $G$ is a homogeneous polynomial with degree $d$. Thus, $G/H(P_i) = G(\alpha_i, 1)$, which is an one-variable polynomial with degree less than $k$.

Therefore, because of this bijection between $\mathcal L_k$ and $\mathcal L((k-1)P_\infty)$, we can rewrite the Reed--Solomon code from Definition \ref{defn:rs_codes} as
\begin{equation}
	RS_q(n, k) = \left\{\big(f(P_1), f(P_2), \dots, f(P_n)\big) \mid f \in \mathcal L((k-1)P_\infty)\right\}.
\end{equation}

@@

\begin{thm}[\cite{Wal00}]
	Let $\mathfrak C, \mathcal P, D$ be as above. Let $g$ denote the genus of $\mathfrak C$. Suppose $D$ satisfies $2g - 2 < \deg D < n$. Then the algebraic geometric code $C(\mathfrak C, \mathcal P, D)$ is a linear code with 
length $n$,
dimension $\deg D + 1 - g$, and 
minimum distance $d \ge n - \deg D$.
\end{thm}

@@

This theorem is an interesting result! We want large $R$ and $\delta$, and these codes yield $k + d \ge n + 1 - g$, i.e., 
	\begin{equation}
		\label{eq:g/n} R + \delta \ge 1 + 1/n - g/n,
	\end{equation}
where $n$ is the number of rational points of the curve $\mathfrak C$ with genus $g$.
Futhermore, all the work done can be applied for algebraic curves in higher dimensional projective spaces.

@@

On equation \eqref{eq:g/n}, we observe that good algebraic geometric codes are generated by curves with a small ratio between $g$ and $n$.
On \cite{TVZ82}, 
%\todo{Write one or two more sentences about this article.}
the authors present a sequence of such curves, with $g/n$ large enough to create a better bound than the Gilbert--Varshamov one.
